---
title: "Midterm Project"
author: "Yuxin Zeng"
date: "2020/11/28"
output: pdf_document
---

```{r setup, include=FALSE}
pacman::p_load("tidyverse","lubridate","forecast","ggplot2","gridExtra","qcc","MASS","arm","blmeco","epiDisplay","faraway","bayesplot")
```

# Data
First read the COVID-19 cases data. Only Chicago residents are included based on the home ZIP Code, as provided by the medical provider, or the address, as provided by the Cook County Medical Examiner. Confirmed cases are counted on the date the test specimen was collected.
See more details at https://healthdata.gov/dataset/covid-19-daily-cases-deaths-and-hospitalizations
```{r include=FALSE}
dt=read.csv("COVID-19_Daily_Cases__Deaths__and_Hospitalizations.csv")
dt$Date=as.Date(dt$Date,"%m/%d/%Y") 
dt=dt[month(dt$Date)<=08,]
```

Since high temperature may inhibit some virus activity. I am curious whether temperature affects the number of confirmed cases per day. I downloaded and organized the temperature data from NDBC website. The air temperature (ATMP) is measured every hour, and I take the temperature at 12 noon as the temperature of that day. The unit is Fahrenheit (F).
See more details at  https://www.ndbc.noaa.gov/station_history.php?station=chii2
```{r include=FALSE}
Mar=data.frame(read.table(file="mar.txt")) 
Apr=data.frame(read.table(file="apr.txt"))
May=data.frame(read.table(file="may.txt"))
Jun=data.frame(read.table(file="jun.txt"))
Jul=data.frame(read.table(file="jul.txt"))
Aug=data.frame(read.table(file="aug.txt"))
temp=data.frame(rbind(Mar,Apr,May,Jun,Jul,Aug))
colnames(temp)=c("YY","MM","DD","hh","mm","WDIR","WSPD","GST","WVHT","DPD","APD","MWD","PRES","ATMP","WTMP","DEWP","VIS","TIDE")
temp=temp[temp$hh=="12"&temp$mm=="0",] 
temp$Date=make_datetime(temp$YY,temp$MM,temp$DD)  
temp=temp[,c("ATMP","Date")]
```

COVID-19 cases data is available from March to October.Temperature data is available from January to August. Merge the two data sets and the new data set contains temperature and cases information from March to August.
```{r include=FALSE}
dt=left_join(temp,dt,by="Date")
which(is.na(dt))
dt=dt[,c(1:5,27:31)]
```

I tried to add weekend and holidays variable to indicate the gathering of people. The "weekend" variable is 0 for weekdays, 1 for weekends. Federal holidays that between March and August are Memorial Day (2020-05-25) and Independence Day (2020-07-04), and both have a 3-day off (includes weekend). Taking the asymptomatic of COVID-19 into account, assume there is a two-week incubation period for each holiday. That is, the "holidays" variable is 2 for dates within two weeks after Memorial Day, 1 for dates within two weeks after Independence Day, 0 for otherwise. (Since Independence Day is in summer vacation, for student groups, there is no big difference on that day compared with other days.)
```{r include=FALSE}
#Weekday
Sys.setlocale("LC_TIME", "English")
dt$weekday=weekdays(as.Date(dt$Date))
dt$weekend=ifelse(dt$weekday=="Saturday" | dt$weekday=="Sunday",1,0)

#Holiday
#Memorial Day 
M=difftime(as.Date("2020-05-25"),as.Date(dt$Date),units="days")  
#Independence Day
I=difftime(as.Date("2020-07-24"),as.Date(dt$Date),units="days")  

dt$holidays=ifelse(0<=I & I<=14,1,ifelse(0<=M & M<=14,2,0))
```

# EDA
```{r echo=FALSE}
p1=ggplot(dt,aes(x=factor(weekend),y=Cases...Total)) + 
  geom_violin(aes(fill=factor(weekend)),color="transparent",alpha=0.5) +
  geom_boxplot(outlier.alpha=0,coef=0,color="gray40",width=0.1) +
  labs(x="Weekdays",y="Number of Cases",title="Number of Cases") +
  scale_fill_discrete(labels=c("Workday","Weekend")) +
  theme(panel.background=element_rect(fill="ivory1"),
        panel.grid.major=element_line(color="gray70",size=0.2),
        panel.grid.minor=element_line(color="gray90",size=0.1),)

p2=ggplot(dt,aes(x=factor(holidays),y=Cases...Total)) + 
  geom_violin(aes(fill=factor(holidays)),color="transparent",alpha=0.5) +
  geom_boxplot(outlier.alpha=0,coef=0,color="gray40",width=0.1) +
  labs(x="Holidays",y="Number of Cases",title="Number of Cases") +
  scale_fill_discrete(labels=c("None","Independence Day 07-04","Memorial Day 05-25")) +
  theme(panel.background=element_rect(fill="ivory1"),
        panel.grid.major=element_line(color="gray70",size=0.2),
        panel.grid.minor=element_line(color="gray90",size=0.1),)

grid.arrange(p1,p2,ncol=1)    
```
Notice the violin plot comparing the weekdays and weekends, the decrease in cases on weekends may probably due to the fact that hospital's inspectors do not work on weekends. So I made an adjustment on "weekend" variable: "weekend_delay" is 1 for Monday, and 0 for otherwise.
```{r include=FALSE}
dt$weekend_delay=ifelse(dt$weekday=="Monday",1,0)
```

# Negative Binomial Model
The dependent variable (number of confirmed cases each day) is a count variable, and the independent variables contain a continuous variable (Air temperature at noon each day) and two categorical variables (weekends and holidays).  
The result of over-dispersion test showed that the data was over-dispersion, so I chose negative binomial model rather than Poisson model. 
The proportion of zero value in the dependent variable is not large, so zero-inflated negative binomial regression is not needed.
```{r include=FALSE}
#Over Dispersion
qcc.overdispersion.test(dt$Cases...Total,type="poisson")

#Zero Inflation
table(dt$Cases...Total=="0")

#Model
fit1=glm.nb(formula=Cases...Total~ATMP+factor(weekend_delay)+factor(holidays),data=dt)
summary(fit1)
```

I used glm.nb function from MASS package to fit the negative binomial model. 
From the summary, air temperature and Memorial Day pass the significance test, indicating that they do affect the number of COVID-19 cases, though the coefficient is not very big. Exclude insignificant variables and fit a new model. The formula is Cases_Total~ATMP+holiday.

```{r include=FALSE}
dt$holiday=ifelse(0<=M & M<=14,1,0)
fit2=glm.nb(formula=Cases...Total~ATMP+holiday,data=dt)
summary(fit2)
```

```{r echo=FALSE}
#Coefficient
exp(coef(fit2))

#Residual Plots
par(mfrow=c(1,2)) 
plot(fitted(fit2),resid(fit2,type="response"),xlab="",ylab="")
title(main="residual plot",xlab="fitted values",ylab="residuals")
binnedplot(fitted(fit2),resid(fit2,type="response"))

#Confidence Interval
confint(fit2)

#Goodness of Fit
poisgof(fit2)
```
The summary shows as the temperature increases, the number of cases decreases slightly. And the number of cases has increased after holidays.
The residuals are mostly clustered on the left in the residual plot. And there are some outliers in the binned residual plot. The 97.5% confidence interval of the coefficients is obtained via confint function. After checking the goodness of fit (poisgof), the p value is around 0.05, so it's not a good fit.

```{r include=FALSE}
fit3=glmer(formula=Cases...Total~ATMP+holiday+(1|(month(dt$Date))),data=dt,family=negative.binomial(theta=1.241))
```
Then I used glmer function to fit a multilevel negative binomial model. I group the data by month, so there are 6 groups in the multilevel negative binomial model. The formula is Cases_Total~ATMP+holiday+(1|(month(Date))). In this sense, the influence of air temperature on the dependent variable goes down a lot because the temperature increases month by month and grouping data by months will weaken the effect of temperature. 
```{r include=FALSE}
summary(fit3)
anova(fit3)

par(mfrow=c(1,2)) 
plot(fitted(fit3),resid(fit3,type="response"),xlab="",ylab="")
title(main="residual plot",xlab="fitted values",ylab="residuals")
binnedplot(fitted(fit2),resid(fit2,type="response"))

dispersion_glmer(fit3) 
```

# ARIMA
```{r include=FALSE}
#Time Series
Number.of.Cases=ts(dt$Cases...Total,frequency=7)
tsdisplay(Number.of.Cases,xlab="Time",ylab="Cases")
fit4=auto.arima(Number.of.Cases)
summary(fit4)
```

```{r echo=FALSE}
#Residual Plot
checkresiduals(fit4)
#Prediction
fore=forecast(fit4)
plot(fore)
lines(fore$fitted,col="steelblue")
```
I fitted a ARIMA(0,1,2)(1,0,0)[7] model. The result of Ljung-Box test shows that the sequence is correlated, not a white noise sequence. The AIC of ARIMA model is smaller than that of negative binomial model.

# Discussion
When counting the number of COVID-19 cases each day, ARIMA model is better than negative binomial regression model, but it is better to use the infectious disease dynamics models like SIR, SEIR, etc.,which can reflect the epidemic law from the aspect of disease transmission mechanism.
The poor fit of the negative binomial regression model is reasonable. Traditional statistical regression models are not suitable for infectious diseases. Compared with the SIR model, they are static, lack geospatiality, and cannot perfectly measure the impact of human activities on the number of cases.
The impact of air temperature and holidays on the number of COVID-19 cases is extremely slight, especially at the beginning of the outbreak, the number is more determined by human activities. If I could collect data on whether people in Chicago wear masks and how often they go out from March to August as independent variables, the model might fit better. 

# Appendix
```{r echo=TRUE}
#Data Preparation
dt=read.csv("COVID-19_Daily_Cases__Deaths__and_Hospitalizations.csv")
dt$Date=as.Date(dt$Date,"%m/%d/%Y") 
dt=dt[month(dt$Date)<=08,]

Mar=data.frame(read.table(file="mar.txt")) 
Apr=data.frame(read.table(file="apr.txt"))
May=data.frame(read.table(file="may.txt"))
Jun=data.frame(read.table(file="jun.txt"))
Jul=data.frame(read.table(file="jul.txt"))
Aug=data.frame(read.table(file="aug.txt"))
temp=data.frame(rbind(Mar,Apr,May,Jun,Jul,Aug))
colnames(temp)=c("YY","MM","DD","hh","mm","WDIR","WSPD","GST","WVHT","DPD","APD","MWD","PRES","ATMP","WTMP","DEWP","VIS","TIDE")
temp=temp[temp$hh=="12"&temp$mm=="0",] 
temp$Date=make_datetime(temp$YY,temp$MM,temp$DD)  
temp=temp[,c("ATMP","Date")]

dt=left_join(temp,dt,by="Date")
which(is.na(dt))
dt=dt[,c(1:5,27:31)]

#Add variable
Sys.setlocale("LC_TIME", "English")
dt$weekday=weekdays(as.Date(dt$Date))
dt$weekend=ifelse(dt$weekday=="Saturday" | dt$weekday=="Sunday",1,0)
dt$weekend_delay=ifelse(dt$weekday=="Monday",1,0)

M=difftime(as.Date("2020-05-25"),as.Date(dt$Date),units="days")  
I=difftime(as.Date("2020-07-24"),as.Date(dt$Date),units="days")  
dt$holidays=ifelse(0<=I & I<=14,1,ifelse(0<=M & M<=14,2,0))

#Negative Binomial Model
#Over Dispersion
qcc.overdispersion.test(dt$Cases...Total,type="poisson")
#Zero Inflation
table(dt$Cases...Total=="0")
#Model
fit1=glm.nb(formula=Cases...Total~ATMP+factor(weekend_delay)+factor(holidays),data=dt)
dt$holiday=ifelse(0<=M & M<=14,1,0)
fit2=glm.nb(formula=Cases...Total~ATMP+holiday,data=dt)

#Multilevel negative binomial
fit3=glmer(formula=Cases...Total~ATMP+holiday+(1|(month(dt$Date))),data=dt,family=negative.binomial(theta=1.241))

summary(fit3)
anova(fit3)

par(mfrow=c(1,2)) 
plot(fitted(fit3),resid(fit3,type="response"),xlab="",ylab="")
title(main="residual plot",xlab="fitted values",ylab="residuals")
binnedplot(fitted(fit2),resid(fit2,type="response"))

dispersion_glmer(fit3) 

#ARIMA
Number.of.Cases=ts(dt$Cases...Total,frequency=7)
tsdisplay(Number.of.Cases,xlab="Time",ylab="Cases")
fit3=auto.arima(Number.of.Cases)
```
